---
title: "Maitland et al. Non-lethal modeling"
author: "Bryan M. Maitland"
date: "14 May 2020"
output:
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r Set Up}
# Load libriaries
library(tidyverse)  # for data import, wrangling, and plotting
library(broom)  # for tidy model objects
library(ggpubr)  # for plotting fxs
library(rstatix)   # pipe friendly stats

# Load custom functions 
source("code/fx_theme_pub.R", chdir = TRUE)
source("code/fx_cleave-by.R", chdir = TRUE)
```

# 1. Data preprocessing
```{r Data Import and Checking}

data_all <- read_csv("data/data_iso_fish_fins-compiled.csv") 
data_fish_groups <- read_csv("data/metadata_fishes.csv")

(df1 <- 
  data_all %>% 
  filter(!taxon_code %in% c("BNT","WAE","STC")) %>%
  droplevels() %>%   
  mutate_if(is.character, as.factor) %>% 
  left_join(data_fish_groups, by = "taxon_code") %>% 
  select(unique_id, site_id, taxon_code, common_name, length_mm, d15N, d13C, CN, d15N_fin, d13C_fin, CN_fin) 
)
```

```{r sample sizes}
# Sample size of dataset
df1 %>% group_by(common_name) %>% tally() %$% sum(n)

# Sample sizes per species
df1 %>% group_by(common_name) %>% tally() %>% arrange(desc(n)) 

```

## Lipid corrections
```{r Check C:N ratios across species}
(df1 %>% 
  group_by(common_name) %>% 
  summarise(mean_CN = mean(CN), 
            sd_CN = sd(CN), 
            mean_CN_fin = mean(CN_fin), 
            sd_CN_fin = sd(CN_fin))
)
```
Most species have relatively low C:N ratios (<4).

```{r Apply lipid correction}
(df1 <- df1 %>% 
  mutate(d13C_lc = if_else(CN > 4.0, 
                               d13C - 3.32 + (0.99 * CN), 
                               d13C),
         d13C_fin_lc = if_else(CN_fin > 4.0, 
                               d13C_fin - 3.32 + (0.99 * CN_fin), 
                               d13C_fin))
)
```

## Make long df
```{r Make Long Data}

# First, select out the muscle data
a <- 
  df1 %>% 
  select(taxon_code, 
        length_mm, 
        d13C, d13C_lc, d15N) %>% 
    mutate(sample_type = "muscle")

# Then select out the fin data
b <- 
  df1 %>% 
  select(taxon_code, 
        length_mm, 
        d13C_fin, d13C_fin_lc, d15N_fin) %>% 
  mutate(sample_type = "fin") %>% 
  # Rename columns to match df a above
  rename(d13C = d13C_fin, d13C_lc = d13C_fin_lc, d15N = d15N_fin)

# Now Stack'em
data_long <- 
  a %>% 
  bind_rows(b) %>%
  left_join(data_fish_groups, by = "taxon_code")
data_long
```


# 2. Data Exploration

## Data summary
```{r Summary table}
data_summary <- df1 %>% 
  # Calculate descriptive statistics
  mutate(oset_C = d13C_fin_lc - d13C_lc, 
         oset_N = d15N_fin - d15N) %>% 
  group_by(taxon_code) %>% 
  summarise(N = n(),
            length_m = mean(length_mm), 
            length_sd = sd(length_mm), 
            cn_fin_m = mean(CN_fin), 
            cn_fin_sd = sd(CN_fin), 
            cn_mus_m = mean(CN), 
            cn_mus_sd = sd(CN),
            oset_C_m = mean(oset_C), 
            oset_C_sd = sd(oset_C), 
            oset_N_m = mean(oset_N), 
            oset_N_sd = sd(oset_N)
            ) %>% 
  # Tidy and wrtie to file
  mutate_if(is.numeric, round, 2) %>% 
  mutate(length_m = round(length_m, 0), 
         length_sd = round(length_sd, 0)) %>% 
  mutate(Length_m = paste0(length_m, " (", length_sd, ")"), 
         CN_fin = paste0(cn_fin_m, " (", cn_fin_sd, ")"), 
         CN_mus = paste0(cn_mus_m, " (", cn_mus_sd, ")"),
         Offset_C = paste0(oset_C_m, " (", oset_C_sd, ")"),
         Offset_N = paste0(oset_N_m, " (", oset_N_sd, ")"),
         ) %>% 
  left_join(data_fish_groups, by = "taxon_code") %>% 
  select(family, taxon, common_name, N, Length_m, CN_fin, CN_mus, Offset_C, Offset_N) %>% 
  arrange(family, common_name) %>% 
  rename(Family = family, Species=taxon, Common=common_name) %>% 
  write_csv("tables/data_summary.csv")
data_summary
```

## Offsets summary
```{r Isotope offsets}
df1 %>% 
  summarise(n = n(), 
            Cfin_mean = mean(d13C_fin), 
            Cfin_sd = sd(d13C_fin), 
            C_mean = mean(d13C), 
            C_sd = sd(d13C),             
            Nfin_mean = mean(d15N_fin), 
            Nfin_sd = sd(d15N_fin), 
            N_mean = mean(d15N), 
            N_sd = sd(d15N)            
            )


df1 %>% 
  mutate(oset_C = d13C_fin_lc - d13C_lc, 
         oset_N = d15N_fin - d15N) %>% 
  #group_by(taxon_code) %>% 
  summarise(n = n(), 
            oset_C_mean = mean(oset_C), 
            oset_C_sd = sd(oset_C), 
            oset_N_mean = mean(oset_N), 
            oset_N_sd = sd(oset_N)) 

```

```{r C:N ratio offsets}

df1 %>% 
  mutate(oset_CN = CN_fin - CN) %>% 
  group_by(taxon_code) %>% 
  summarise(n = n(), 
            oset_CN_mean = mean(oset_CN), 
            oset_CN_sd = sd(oset_CN)) %>% 
  summarise(mean_CN_oset = mean(oset_CN_mean), 
            sd_CN_offset = sd(oset_CN_mean))

df1 %>% 
  mutate(oset_CN = CN_fin - CN) %>% 
  group_by(taxon_code) %>% 
  summarise(n = n(), 
            oset_CN_mean = mean(oset_CN), 
            oset_CN_sd = sd(oset_CN))

```

## Boxplots
```{r fig - Boxplots and outliers}

(outliers.p <- 
  ggpubr::ggarrange(

    # Carbon
    data_long %>% 
      ggplot(aes(x = sample_type, y = d13C_lc)) + 
      geom_boxplot(outlier.shape = 21) +
      scale_y_continuous(limits = c(-34,-22), breaks = seq(-34, -22, 2)) + 
      theme_Publication() + 
      labs(x = "Tissue type",
           y = expression(~{delta}^13*C['lipid corrected']~'(\u2030)'), 
           title = "(A) Carbon values"),

    # Nitrogen
    data_long %>% 
      ggplot(aes(x = sample_type, y = d15N)) + 
      geom_boxplot(outlier.shape = 21) +
      scale_y_continuous(limits = c(6, 16)) + 
      theme_Publication() + 
      labs(x = "Tissue type",
           y = expression(~{delta}^15*N~'(\u2030)'), 
           title = "(B) Nitrogen values")
    )
 )

# Save to ffile
ggsave("figs/boxplot_by-tissue.pdf", plot = outliers.p, device = cairo_pdf, 
       height = 4, width = 8)
```

### Offset stats

```{r test for differences between tissues}
hist(df1$d15N_fin)
hist(df1$d15N)

shapiro_test(log10(df1$d15N_fin))
shapiro_test(log10(df1$d15N))

shapiro_test(sqrt(df1$d15N_fin))
shapiro_test(sqrt(df1$d15N))

hist(sqrt(df1$d15N_fin))
hist(sqrt(df1$d15N))



t.test(df1$d13C_fin_lc, df1$d13C_lc, paired = TRUE)
t.test(df1$d15N_fin, df1$d15N, paired = TRUE)

lsr::cohensD(df1$d13C_fin_lc, df1$d13C_lc, method = "paired")
lsr::cohensD(df1$d15N_fin, df1$d15N, method = "paired")


# wilcox.test(df1$d13C_fin_lc, df1$d13C_lc, paired = TRUE)
# wilcox.test(df1$d15N_fin, df1$d15N, paired = TRUE)

```

# 3. Model Assumptions

## Independence of observations
Because we only have one independent variable (fins isotope values) and one dependent variable (muscle isotope values), we don’t need to test for any hidden relationships among variables.

## Normality
```{r muscle carbon}
ggpubr::ggarrange(
  ggpubr::ggdensity(df1$d13C_lc, 
                    main = "Density Plot", xlab = "d13C_lc"),
  ggpubr::ggqqplot(df1$d13C_lc, 
                   main = "Q-Q Plot", xlab = "d13C_lc"), ncol = 2)

shapiro.test(df1$d13C_lc)
```

```{r fin carbon}
ggpubr::ggarrange(
  ggpubr::ggdensity(df1$d13C_fin_lc, 
                    main = "Density Plot", xlab = "d13C_fin_lc"),
  ggpubr::ggqqplot(df1$d13C_fin_lc, 
                   main = "Q-Q Plot", xlab = "d13C_fin_lc"), ncol = 2)

shapiro.test(df1$d13C_fin_lc)
```

```{r muscle nitrogen}
ggpubr::ggarrange(
  ggpubr::ggdensity(df1$d15N, 
                    main = "Density Plot", xlab = "d15N"),
  ggpubr::ggqqplot(df1$d15N, 
                   main = "Q-Q Plot", xlab = "d15N"), ncol = 2)

shapiro.test(df1$d15N)
```

```{r fin nitrogen}
ggpubr::ggarrange(
  ggpubr::ggdensity(df1$d15N_fin, 
                    main = "Density Plot", xlab = "d15N_fin"),
  ggpubr::ggqqplot(df1$d15N_fin, 
                   main = "Q-Q Plot", xlab = "d15N_fin"),ncol = 2)

shapiro.test(df1$d15N_fin)
```

Carbon is normallly distributed, nitorgen not, which is skewed. 

## Linearity
```{r carbon scatterplot}
df1 %>% 
  ggplot(aes(x = d13C_fin_lc, y = d13C_lc)) + 
  geom_point() + 
  geom_abline(linetype = "dashed", slope = 1, intercept = 0, color = "grey")

```

```{r nitrogen scatterplot}
df1 %>% 
  ggplot(aes(x = d15N_fin, y = d15N)) + 
  geom_point() + 
  geom_abline(linetype = "dashed", slope = 1, intercept = 0, color = "grey")
```
Muscle-fin isotope relationships are linear for both carbon and nitrogen. 

## Ontogenetic shifts

```{r fig - ontogentic effects}
ratios <- df1 %>% 
  mutate(ratio_C = d13C_fin_lc / d13C_lc, 
         ratio_N = d15N_fin / d15N) 

# Carbon----

# # Pearson correlation
# df1 %>% 
#   mutate(ratio_C = d13C_fin / d13C) %$%
#   cor.test(ratio_C, length_mm, method = "pearson") %>% 
#   tidy()
# 
# # Check slope values
# df1 %>% 
#   mutate(ratio_C = d13C_fin / d13C) %>% 
#   lm(ratio_C ~ length_mm + taxon_code, data = .) %>% 
#   tidy()

# Plot
c_fin_ratio_size <- 
  df1 %>% 
  mutate(ratio_C = d13C_fin_lc / d13C_lc) %>% 
  ggplot(aes(x = length_mm, y = ratio_C)) +
  geom_hline(yintercept = 1) +
  geom_point(color = "black", fill = "gray", shape = 21) +
  stat_cor(method = "pearson", label.x = 200, label.y = 1.1, size = 3)+ 
  scale_x_continuous(limits = c(0,500)) + 
  scale_y_continuous(limits = c(0.78, 1.1)) + 
  theme_Publication() +
  labs(x = "Total length (mm)",
       y = expression(~{delta}^13*C["fin"]~":"~{delta}^13*C["muscle"]), 
       title = "(A) Carbon values")


# Nitrogen -----

# # Plot
# # Pearson correlation
# df1 %>% 
#   mutate(ratio_N = d15N_fin / d15N) %$%
#   cor.test(ratio_N, length_mm, method = "pearson") %>% 
#   tidy()
# 
# # Check slope values
# df1 %>% 
#   mutate(ratio_N = d15N_fin / d15N) %$%
#   lm(ratio_N ~ length_mm + taxon_code, data = .) %>% 
#   tidy()

n_fin_ratio_size <- 
  df1 %>% 
  mutate(ratio_N = d15N_fin / d15N) %>% 
  ggplot(aes(x = length_mm, y = ratio_N)) +
  geom_hline(yintercept = 1) +
  geom_point(color = "black", fill = "gray", shape = 21) +
  stat_cor(method = "pearson", label.x = 200, label.y = 1.1, size = 3) + 
  scale_x_continuous(limits = c(0,500)) + 
  scale_y_continuous(limits = c(0.78, 1.1)) + 
  theme_Publication() +
  labs(x = "Total length (mm)",
       y = expression(~{delta}^15*N["fin"]~":"~{delta}^15*N["muscle"]), 
       title = "(B) Nitrogen values")

( fin_mus_ratio_cor <- ggpubr::ggarrange(c_fin_ratio_size, n_fin_ratio_size) )

ggsave("figs/fin-mus-ratio-corr.pdf", plot = fin_mus_ratio_cor, device = cairo_pdf,
       height = 4, width = 8)

```


# 4. Linear modeling
```{r set seed for reproducibility}
set.seed(09222017)
```

```{r Split Data for Modeling}
# Training dataset
data_train <- df1 %>% 
  group_by(taxon_code) %>% 
  sample_frac(0.8) %>% 
  ungroup() %>% 
  arrange(common_name)

# Test dataset
data_test <- anti_join(df1, data_train, by = 'unique_id') %>% arrange(common_name)

# Sample sizes 

data_train %>% 
  group_by(common_name) %>% 
  tally() %>% 
  summarise(mean = sum(n))

data_test %>% 
  group_by(common_name) %>% 
  tally() %>% 
  summarise(mean = sum(n))

(spp_samp_size <- 
  data_train %>% 
  group_by(common_name) %>% 
  tally())

```

## Global Models
### Fit models
```{r Fit Global Models}
fit.global.C <- lm(d13C_lc ~ d13C_fin_lc, data = data_train)
fit.global.N <- lm(d15N ~ d15N_fin, data = data_train)
```

### Model summary
```{r Global Model Summary}
# summary(fit.global.C)
glance(fit.global.C)
# par(mfrow=c(2,2))
# plot(fit.global.C)
# par(mfrow=c(1,1))

# summary(fit.global.N)
glance(fit.global.N)
# par(mfrow=c(2,2))
# plot(fit.global.N)
# par(mfrow=c(1,1))

#confint_tidy(fit.global.N, parm = "d15N_fin")
```

#### Cook's distcance
Cook’s distance is a measure computed with respect to a given regression model and therefore is impacted only by the X variables included in the model. But, what does cook’s distance mean? It computes the influence exerted by each data point (row) on the predicted outcome.

```{r Cook's distance}
cooksd_C <- cooks.distance(fit.global.C)
cooksd_C_line <- 4 * mean(cooksd_C, na.rm = TRUE)
# plot(cooksd_C, pch="*", cex=2, main="C Influential Obs by Cooks distance")  # plot cook's distance
# abline(h = 4*mean(cooksd_C, na.rm=T), col="red")  # add cutoff line
# text(x=1:length(cooksd_C)+1, y=cooksd_C, 
#      labels=ifelse(cooksd_C>4*mean(cooksd_C, na.rm=T), 
#                    names(cooksd_C), ""), col = "red")  # add labels

cooksd_C %>% 
  enframe() %>% 
  ggplot(aes(x = as.integer(name), y = value)) + 
  geom_hline(yintercept = cooksd_C_line, linetype = "dashed") +
  geom_point() + 
  geom_point(data = subset(enframe(cooksd_C), value > cooksd_C_line), 
             color = "red") + 
  labs(x = "Sample ID", y = "Cook's Distance", 
       title = "(A) Multispecies carbon regression model") + 
  theme_Publication() ->
  cook_c
cook_c


cooksd_N <- cooks.distance(fit.global.N)
cooksd_N_line <- 4 * mean(cooksd_N, na.rm = TRUE)
# plot(cooksd_N, pch="*", cex=2, main="N Influential Obs by Cooks distance")  # plot cook's distance
# abline(h = 4*mean(cooksd_N, na.rm=T), col="red")  # add cutoff line
# text(x=1:length(cooksd_N)+1, y=cooksd_N, 
#      labels=ifelse(cooksd_N>4*mean(cooksd_N, na.rm=T), 
#                    names(cooksd_N), ""), col = "red")  # add labels

cooksd_N %>% 
  enframe() %>% 
  ggplot(aes(x = as.integer(name), y = value)) + 
  geom_hline(yintercept = cooksd_N_line, linetype = "dashed") +
  geom_point() + 
  geom_point(data = subset(enframe(cooksd_N), value > cooksd_N_line), 
             color = "red") +   
  labs(x = "Sample ID", y = "Cook's Distance", 
       title = "(B) Multispecies nitorgen regression model") + 
   theme_Publication() ->
  cook_n
cook_n

ggpubr::ggarrange(cook_c, cook_n, nrow = 2) -> cook_panel
cook_panel

ggsave("figs/cook_panel.pdf", plot = cook_panel, device = cairo_pdf, 
       height = 5, width = 8)
```

Now lets find out the influential rows from the original data. If you extract and examine each influential row 1-by-1 (from below output), you will be able to reason out why that row turned out influential. It is likely that one of the X variables included in the model had extreme values.

```{r influential obs.}
# influential row numbers
influential_C <- as.numeric(names(cooksd_C)[(cooksd_C > 4*mean(cooksd_C, na.rm=T))])

# influential observations
data_train[influential_C, ]


# influential row numbers
influential_N <- as.numeric(names(cooksd_N)[(cooksd_N > 4*mean(cooksd_N, na.rm=T))])

# influential observations
data_train[influential_N, ]
```

### Plot global models
```{r Plot global models}
# Carbon ----
# Set custom labels
format_pval <- function(pval){
  pval <- scales::pvalue(pval, accuracy= 0.001, add_p = TRUE)
  gsub(pattern = "(=|<)", replacement = " \\1 ", x = pval)
}
# model_label <- c("italic(P) < 0.001", 
#                  paste("Adj~italic(R)^2 ==", signif(summary(fit.global.C)$adj.r.squared, 2)))
# Set coords
min(data_train$d13C_fin_lc)
max(data_train$d13C_fin_lc)
min(data_train$d13C_lc)
max(data_train$d13C_lc)

coord_x <- min(data_train$d13C_fin_lc)
coord_y <- max(data_train$d13C_lc)

# Get model fiT and 95% CIs and bind to raw data
fit_pred <- fit.global.C %>% 
  predict(newdata = data_train, interval = 'confidence') %>% 
  as_tibble() %>% 
  bind_cols(data_train)

# Plot
p.lfit.global.Carbon <- 
  fit_pred %>% 
  left_join(data_fish_groups, by = "common_name") %>% 
  ggplot(aes(x = d13C_fin_lc, y = d13C_lc)) + 
  geom_abline(linetype = "dashed", slope = 1, intercept = 0, color = "grey45") +
  # geom_abline(linetype = "dashed", slope = 0.82, intercept = -5.89, color = "red") +
  # geom_abline(linetype = "dashed", slope = 0.89, intercept = -3.27, color = "blue") +
  geom_point(color = "black", fill = "grey90", shape = 21, size = 2, alpha = 0.75) +
  geom_ribbon(aes(x = d13C_fin_lc, ymin = lwr, ymax = upr), alpha = 0.3, color = NA) +
  geom_line(aes(x = d13C_fin_lc, y = fit), size = 1) +
  scale_x_continuous(limits = c(-31,-22), breaks = seq(-30,-22,2)) +
  scale_y_continuous(limits = c(-32.5,-21.7), breaks = seq(-32,-22,2)) +
  labs(x = expression(~{delta}^13*C["fin"]~'\u2030'), 
       y = expression(~{delta}^13*C["muscle"]~'\u2030'), 
       shape = "Family",
       fill = "",
       title = "(A) Multi-species carbon model") +
  stat_cor(label.y = -22.5, aes(label = paste(..rr.label.., format_pval(..p..), sep = "*`,`~")), size = 3) +
  stat_regline_equation(label.y = -24, size = 3)+
  annotate(geom = "text", x = -24.5, y = -23, label = "1:1 line", color = "grey45", family = "Times") +
  # annotate(geom = "text", x = coord_x, y = c(-22, -23), 
  #            hjust = 0, label = model_label, parse = TRUE, family = "Times") +
  theme_Publication() + 
  theme(plot.margin = unit(c(1,1,1,1),"mm"))

p.lfit.global.Carbon

# ggsave(filename = "figs/scatterplot-global-carbon.pdf",
#        plot = p.lfit.global.Carbon, device = cairo_pdf,
#        units = "in", width = 8, height = 6)

# Nitrogen ----
#Set custom labels
model_label <- c("italic(P) < 0.001", 
                 paste("Adj~italic(R)^2 ==", signif(summary(fit.global.N)$adj.r.squared, 2)))
#Set coords
min(data_train$d15N_fin)
max(data_train$d15N_fin)
min(data_train$d15N)
max(data_train$d15N)

coord_x <- min(data_train$d15N_fin)
coord_y <- max(data_train$d15N)

# Get model fiT and 95% CIs and bind to raw data
fit_pred <- fit.global.N %>% 
  predict(newdata = data_train, interval = 'confidence') %>% 
  as_tibble() %>% 
  bind_cols(data_train)

#Plot
p.lfit.global.Nitorgen <- 
  fit_pred %>% 
  left_join(data_fish_groups, by = "common_name") %>% 
  ggplot(aes(x = d15N_fin, y = d15N)) + 
  geom_abline(linetype = "dashed", slope = 1, intercept = 0, color = "grey45") +
  # geom_abline(linetype = "dashed", slope = 1.01, intercept = 0.74, color = "red") +
  # geom_abline(linetype = "dashed", slope = 0.81, intercept = 1.73, color = "blue") +
  geom_point(color = "black", fill = "grey90", shape = 21, size = 2, alpha = 0.75) +
  geom_ribbon(aes(x = d15N_fin, ymin = lwr, ymax = upr), alpha = 0.3, color = NA) +
  geom_line(aes(x = d15N_fin, y = fit), size = 1) +
  scale_x_continuous(limits = c(5,16), breaks = seq(6,16,2)) +
  scale_y_continuous(limits = c(6,16), breaks = seq(6,16,2)) +
  labs(x = expression(~{delta}^15*N["fin"]~'\u2030'), 
       y = expression(~{delta}^15*N["muscle"]~'\u2030'), 
       shape = "Family",
       fill = "",
       title = "(B) Multi-species nitrogen model") +
  stat_cor(label.y = 15.5, aes(label = paste(..rr.label.., format_pval(..p..), sep = "~`,`~")), size = 3) +
  stat_regline_equation(label.y = 14, size = 3) +
  annotate(geom = "text", x = 14.5, y = 13, label = "1:1 line", color = "grey45", family = "Times") +
  # annotate(geom = "text", x = coord_x, y = c(15.5, 14.5), 
  #            hjust = 0, label = model_label, parse = TRUE, family = "Times") +
  theme_Publication() + 
  theme(plot.margin = unit(c(1,1,1,1),"mm"))

p.lfit.global.Nitorgen

# ggsave(filename = "figs/scatterplot-global-nitrogen.pdf",
#        plot = p.lfit.global.Nitorgen, device = cairo_pdf,
#        units = "in", width = 8, height = 6)

# Panel plot ----
p.global.panel <- 
  ggpubr::ggarrange(p.lfit.global.Carbon, p.lfit.global.Nitorgen,
                    nrow = 1, ncol = 2)
p.global.panel

ggsave(filename = "figs/fig1.pdf", 
       plot = p.global.panel, device = cairo_pdf,
       units = "in", width = 8.5, height = 4)
```

## Species-Specific Models
### Plot species panels
```{r species panel carbon}
# Custom formatting fucntion
format_pval <- function(pval){
  pval <- scales::pvalue(pval, accuracy= 0.001, add_p = TRUE)
  gsub(pattern = "(=|<)", replacement = " \\1 ", x = pval)
}

p.spp.panel.carbon <- 
  data_train %>% 
  ggplot(aes(x = d13C_fin_lc, y = d13C_lc)) + 
  geom_point(data=data_train[,c("d13C_lc","d13C_fin_lc","unique_id")], 
             aes(x=d13C_fin_lc, y=d13C_lc, group=unique_id), 
             size = 1, colour="grey85") +
  geom_point(size = 1, color = "black", fill = "grey90", shape = 21) +
  stat_cor(label.y = -22.5, aes(label = paste(..rr.label.., format_pval(..p..), sep = "*`,`~")), size = 3) +
  stat_regline_equation(label.y = -24, size = 3)+
  geom_smooth(aes(group = common_name), method = "lm", col = "black", se = FALSE) +
  geom_abline(linetype = "dashed", slope = 1, intercept = 0, color = "black") +
  lemon::facet_rep_wrap(~common_name, nrow = 3) + 
  labs(x = expression(~{delta}^13*C["fin"]~'\u2030'),
       y = expression(~{delta}^13*C["muscle"]~'\u2030')) +
  scale_x_continuous(limits = c(-30.7,-21.7), breaks = seq(-30,-22,4)) +
  scale_y_continuous(limits = c(-32.5,-22), breaks = seq(-32,-22,4)) +
  theme_Publication() 
p.spp.panel.carbon

ggsave(filename = "figs/fig2.pdf", 
       plot = p.spp.panel.carbon, device = cairo_pdf, 
       units = "in", width = 10, height = 6)

```

```{r species panel nitrgoen}
p.spp.panel.nitorgen <- 
  data_train %>% 
  ggplot(aes(x = d15N_fin, y = d15N)) + 
  geom_point(data = data_train[,c("d15N","d15N_fin","unique_id")], 
             aes(x = d15N_fin, y = d15N, group = unique_id), 
             size = 1, colour = "grey85") +
  geom_point(size = 1, color = "black", fill = "grey90", shape = 21) +
  stat_cor(label.y = 15.5, aes(label = paste(..rr.label.., format_pval(..p..), sep = "~`,`~")), size = 3) +
  stat_regline_equation(label.y = 14, size = 3) +
  geom_smooth(aes(group = common_name), method = "lm", col = "black", se = FALSE) +
  geom_abline(linetype = "dashed", slope = 1, intercept = 0, color = "black") +
  lemon::facet_rep_wrap(~common_name, nrow = 3) + 
  labs(x = expression(~{delta}^15*N["fin"]~'\u2030'),
       y = expression(~{delta}^15*N["muscle"]~'\u2030')) +
  scale_x_continuous(limits = c(5.7,16), breaks = seq(6,16,4)) +
  scale_y_continuous(limits = c(6,16), breaks = seq(6,16,4))  +
  theme_Publication()
p.spp.panel.nitorgen

ggsave(filename = "figs/fig3.pdf", 
       plot = p.spp.panel.nitorgen, device = cairo_pdf, 
       units = "in", width = 10, height = 6)
```

### Fit species-specific models
```{r Fit species-specific models}

# create first dataset - 70% for running the model
data_train_nested <- 
  data_train %>% 
  select(d15N, d15N_fin, d13C_lc, d13C_fin_lc, common_name) %>%
  group_by(common_name) %>% 
  nest() %>%
  rename(myorigdata = data)

# create second dataset - 20% for making predictions
data_test_nested <- 
  data_test %>% 
  select(d15N, d15N_fin, d13C_lc, d13C_fin_lc, common_name) %>%
  group_by(common_name) %>% 
  nest() %>% 
  rename(mynewdata = data)

# Do the regressions
regressions <- 
  data_train_nested %>% 
  mutate(
    fit_C = map(.x = myorigdata, .f = ~ lm(d13C_lc ~ d13C_fin_lc, data = .)),
    fit_N = map(.x = myorigdata, .f = ~ lm(d15N ~ d15N_fin, data = .)),

    ci_C = map(fit_C, function(df) confint_tidy(df, parm = "d13C_fin_lc")),
    ci_N = map(fit_N, function(df) confint_tidy(df, parm = "d15N_fin")),

    tidy_C = map(fit_C, tidy),
    tidy_N = map(fit_N, tidy),

    glance_C = map(fit_C, glance),
    glance_N = map(fit_N, glance),

    augment_C = map(fit_C, augment),
    augment_N = map(fit_N, augment)
    ) %>% 
  ungroup()
regressions

```

### Examine residuals
```{r Homoscedasticity}
regressions %>% unnest(augment_C) %>% 
  ggplot(aes(d13C_fin_lc, .resid)) +
    geom_point(alpha = 1 / 3) + 
    geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~common_name)

regressions %>% unnest(augment_N) %>% 
  ggplot(aes(d15N_fin, .resid)) +
    geom_point(alpha = 1 / 3) + 
    geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~common_name) 
```

### Predictions
```{r Predict dataframe}
predictions_spp <- 
  regressions %>% 
  select(common_name, fit_C, fit_N) %>% 
  full_join(data_test_nested, by = "common_name") %>% 
  mutate(pred_spp_C = map2(fit_C, mynewdata, predict),
         pred_spp_N = map2(fit_N, mynewdata, predict)) %>% 
  select(common_name, mynewdata, pred_spp_C, pred_spp_N) %>%
  unnest(c(mynewdata, pred_spp_C, pred_spp_N)) %>% 
  ungroup()
predictions_spp
```

### Calculate error

```{r Model fit summary - R2 and p-vals}
spp_model_summaries <- 
    bind_rows(
      regressions %>% 
        unnest(glance_C, .drop = TRUE) %>% 
        select(common_name, r.squared, p.value) %>%  
        mutate(isotope = "carbon") %>% 
        select(common_name, isotope, r.squared, p.value),
      regressions %>% 
        unnest(glance_N, .drop = TRUE) %>% 
        select(common_name, r.squared, p.value) %>%  
        mutate(isotope = "nitrogen") %>% 
        select(common_name, isotope, r.squared, p.value), 
      bind_rows(
        glance(fit.global.C) %>% 
          select(r.squared, p.value) %>% 
          mutate(common_name = "Global", isotope = "carbon") %>% 
         select(common_name, isotope, r.squared, p.value),
       glance(fit.global.N) %>% 
         select(r.squared, p.value) %>% 
         mutate(common_name = "Global", isotope = "nitrogen") %>% 
         select(common_name, isotope, r.squared, p.value))) %>% 
  mutate(common_name = factor(common_name, levels = c(
      "Global","Bigmouth Shiner","Brassy Minnow","Common Shiner","Creek Chub",
      "Fathead Minnow","Freshwater Drum", "Hornyhead Chub",
      "Johnny Darter","Longnose Dace","Longnose Sucker",
      "Red Shiner", "Sand Shiner","Shorthead Redhorse",
      "Smallmouth Bass","White Sucker"))) %>%
  arrange(isotope,common_name) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = if_else(p.value<0.001,paste("<0.001"),paste(p.value))) %>% 
  print(n= Inf)

spp_model_summaries %>%
  ggplot(aes(x = common_name, y = r.squared, fill = isotope)) +
  geom_point(shape = 21, size = 4) +
  coord_flip() +
  geom_hline(linetype = "dashed", yintercept = 0.8, color = "grey") +
  scale_fill_grey() +
  #scale_y_continuous(limits = c(0.3,1), breaks = seq(0,1,0.1)) +
  labs(x = "Fish Species", y = expression(italic(R)^2),
       title = expression("Comparing Gobal and Species-specific Models")) +
  theme_classic(base_size = 16) +
  theme(axis.text.x = element_text(face="bold", size = 12))

spp_model_summaries %>%
  ggplot(aes(x = common_name, y = p.value, fill = isotope)) +
  geom_point(shape = 21, size = 4) +
  coord_flip() +
  #geom_hline(linetype = "dashed", yintercept = 0.75, color = "grey") +
  scale_fill_grey() +
  #scale_y_continuous(limits = c(0.3,1), breaks = seq(0,1,0.1)) +
  labs(x = "Fish Species", y = "p-val",
       title = expression("Comparing Gobal and Species-specific Models")) +
  theme_classic(base_size = 16) +
  theme(axis.text.x = element_text(face="bold", size = 12))

```

```{r Table of model fits}
df_lm_summary <-
  spp_model_summaries %>% 
  
  left_join(
    bind_rows(
      regressions %>% 
        group_by(common_name) %>% 
        unnest(tidy_C) %>% 
        select(term, estimate) %>% 
        spread(term, estimate) %>% 
        mutate(isotope = "carbon") %>% 
        rename(intc = "(Intercept)", coef = d13C_fin_lc) %>% 
        select(common_name, isotope, intc, coef),
      regressions %>% 
        group_by(common_name) %>% 
        unnest(tidy_N) %>% 
        select(term, estimate) %>% 
        spread(term, estimate) %>% 
        mutate(isotope = "nitrogen") %>% 
        rename(intc = "(Intercept)", coef = d15N_fin) %>% 
        select(common_name, isotope, intc, coef),
      bind_rows(
        tidy(fit.global.C) %>% 
          select(term, estimate) %>% 
          spread(term, estimate) %>% 
          mutate(common_name = "Global", isotope = "carbon") %>% 
          rename(intc = "(Intercept)", coef = d13C_fin_lc) %>% 
          select(common_name, isotope, intc, coef),
        tidy(fit.global.N) %>% 
          select(term, estimate) %>% 
          spread(term, estimate) %>% 
          mutate(common_name = "Global", isotope = "nitrogen") %>% 
          rename(intc = "(Intercept)", coef = d15N_fin) %>% 
          select(common_name, isotope, intc, coef))),
    by = c("common_name","isotope")) %>%
  
  cleave_by(isotope) %>% 
  bind_cols() %>% 
  select(common_name,
         R2_C=r.squared,intc_C=intc,coef_C=coef,Pval_C=p.value, 
         R2_N=r.squared1,intc_N=intc1,coef_N=coef1,Pval_N=p.value1) %>% 
  left_join(
    bind_rows(
      regressions %>% 
        unnest(ci_C, .drop = TRUE) %>% 
        select(common_name, conf.low.C=conf.low, conf.high.C=conf.high) %>% 
        left_join(regressions %>% 
                    unnest(ci_N, .drop = TRUE) %>% 
                    select(common_name, conf.low.N=conf.low, conf.high.N=conf.high),
                  by = "common_name"),
      confint_tidy(fit.global.C, parm = "d13C_fin_lc") %>% 
      mutate(common_name="Global") %>% 
      select(common_name, conf.low.C=conf.low, conf.high.C=conf.high) %>% 
      left_join(confint_tidy(fit.global.N, parm = "d15N_fin") %>% 
                  mutate(common_name="Global") %>% 
                  select(common_name, conf.low.N=conf.low, conf.high.N=conf.high), 
                by="common_name"))) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  mutate(coef_C = paste0(coef_C," (",conf.low.C,"-",conf.high.C,")"), 
         coef_N = paste0(coef_N," (",conf.low.N,"-",conf.high.N,")")) %>% 
  select(-conf.low.C, -conf.low.N, -conf.high.C, -conf.high.N) %>% 
  left_join(spp_samp_size, by = "common_name") %>% 
  left_join(data_fish_groups, by = "common_name") %>% 
  mutate(n = if_else(common_name=="Global", nrow(data_train), n)) %>% 
  arrange(family) %>%   
  select(common_name, n, R2_C, intc_C, coef_C, Pval_C, R2_N, intc_N, coef_N, Pval_N) %>% 
  write_csv("tables/lm_summary.csv")

df_lm_summary

```

# 5. Testing linear models 
```{r Calculate error for each isotope and model by species}
# Calculate error for each isotope
error_d13C <-  predictions_spp %>% 
  mutate(pred_global = predict(fit.global.C, data_test),
         pred_Hette = 0.82 * d13C_fin_lc - 5.89, 
         pred_Jardine  = 0.89 * d13C_fin_lc - 3.27, 
         
         species = abs(d13C_lc - pred_spp_C),
         global = abs(d13C_lc - pred_global),
         hette = abs(d13C_lc - pred_Hette), 
         jardine = abs(d13C_lc - pred_Jardine)) %>% 
  select(common_name, species, global, hette, jardine) %>%
  
  #Now summarize data
  group_by(common_name) %>% 
  summarise_all(list(mean = mean, sd = sd)) %>% 
  gather(key = "key", value = "value", -common_name) %>% 
  separate(key, into = c("model","statistic"), sep = "_") %>% 
  spread(key = statistic, value = value) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(model = factor(model, levels = c("species","global", "hette","jardine"))) %>% 
  mutate(isotope = "Carbon") 
error_d13C

error_d15N <- 
  predictions_spp %>% 
  mutate(pred_global = predict(fit.global.N, data_test),
         pred_Hette = 1.01 * d15N_fin + 0.74, 
         pred_Jardine  = 0.81 * d15N_fin + 1.73, 
         
         species = abs(d15N - pred_spp_N),
         global = abs(d15N - pred_global),
         hette = abs(d15N - pred_Hette), 
         jardine = abs(d15N - pred_Jardine)) %>% 
  select(common_name, species, global, hette, jardine) %>%
  
  #Now summarize data
  group_by(common_name) %>% 
  summarise_all(list(mean = mean, sd = sd)) %>% 
  gather(key = "key", value = "value", -common_name) %>% 
  separate(key, into = c("model","statistic"), sep = "_") %>% 
  spread(key = statistic, value = value) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(model = factor(model, levels = c("species","global", "hette","jardine"))) %>% 
  mutate(isotope = "Nitrogen")
error_d15N

error_df <- bind_rows(error_d13C, error_d15N)
error_df
```

```{r plot model error by species}

labels <- c(Carbon = "(A) Carbon error", Nitrogen = "(B) Nitrogen error")

p.error.panel <- 
  error_df %>% 
  left_join(data_fish_groups, by = "common_name") %>% 
  
  ggplot(aes(x = factor(common_name, levels = rev(levels(factor(common_name)))), 
             y = mean, group = model, color = model, fill = model)) +

  geom_rect(xmin = 1.5, xmax = 2.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 3.5, xmax = 4.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 5.5, xmax = 6.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 7.5, xmax = 8.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 9.5, xmax = 10.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 11.5, xmax = 12.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 13.5, xmax = 14.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +

  geom_abline(linetype = "dashed", slope = 0, intercept = 0.1) +
  geom_pointrange(data = subset(error_df, model %in% c("species","global","hette","jardine")),
                  aes(ymin = mean - sd, ymax = mean + sd),
                  position = position_dodge(width = 0.5),na.rm = TRUE) +
  geom_point(data = subset(error_df, model %in% c("species","global","hette","jardine")), 
             aes(x = factor(common_name, levels = rev(levels(factor(common_name)))), y = mean),
             position = position_dodge(width = 0.5), size = 3, color = "black", shape = 21, na.rm = TRUE) +
  geom_line(data = subset(error_df, model %in% c("species","global","hette","jardine")), 
            position = position_dodge(width = 0.5), na.rm = TRUE) + 
  coord_flip() +
  scale_y_continuous(limits = c(0,2)) +
  scale_colour_manual(values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF"), guide=FALSE) +
  scale_fill_manual(values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF"),
                      labels = c("Species specific", "Multi-species", "European","Australian")) +  
  lemon::facet_rep_wrap(~isotope, labeller=labeller(isotope = labels), 
                        nrow = 1) + 
  labs(x = "", 
       y = expression('Error '~{delta}^x*x~'\u2030'), 
       fill = "Conversion\nModel") +
  theme_Publication() + 
  theme(strip.text=element_text(hjust = 0))

p.error.panel

ggsave(filename = "figs/fig5.pdf",
       plot = p.error.panel, device = cairo_pdf,
       units = "in", width = 10, height = 6)
```

```{r plot model error}
error_d13C_spp <-
  error_d13C %>% 
  filter(isotope=="Carbon") %>% 
  left_join(data_fish_groups, by = "common_name") %>% 
  
  ggplot(aes(x = factor(common_name, levels = rev(levels(factor(common_name)))), 
             y = mean, group = model, color = model, fill = model)) +

  geom_rect(xmin = 1.5, xmax = 2.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 3.5, xmax = 4.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 5.5, xmax = 6.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 7.5, xmax = 8.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 9.5, xmax = 10.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 11.5, xmax = 12.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 13.5, xmax = 14.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +

  geom_abline(linetype = "dashed", slope = 0, intercept = 0.08) +
  geom_pointrange(data = subset(error_d13C, model %in% c("species","global","hette","jardine")),
                  aes(ymin = mean - sd, ymax = mean + sd),
                  position = position_dodge(width = 0.5),na.rm = TRUE) +
  geom_point(data = subset(error_d13C, model %in% c("species","global","hette","jardine")), 
             aes(x = factor(common_name, levels = rev(levels(factor(common_name)))), y = mean),
             position = position_dodge(width = 0.5), size = 3, color = "black", shape = 21, na.rm = TRUE) +
  geom_line(data = subset(error_d13C, model %in% c("species","global","hette","jardine")), 
            position = position_dodge(width = 0.5), na.rm = TRUE) + 
  coord_flip() +
  scale_y_continuous(limits = c(0,2)) +
  scale_colour_manual(values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF"), guide=FALSE) +
  scale_fill_manual(values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF"),
                      labels = c("Species specific", "Multi-species", "European","Australian")) +  
  labs(x = "", 
       y = "", 
       fill = "Conversion\nModel", 
       title = expression('(A)'~{delta}^13*C~'Error')) +
  theme_Publication() 
error_d13C_spp

error_d15N_spp <-
  error_d15N %>% 
  filter(isotope=="Nitrogen") %>% 
  left_join(data_fish_groups, by = "common_name") %>% 
  
  ggplot(aes(x = factor(common_name, levels = rev(levels(factor(common_name)))), 
             y = mean, group = model, color = model, fill = model)) +

  geom_rect(xmin = 1.5, xmax = 2.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 3.5, xmax = 4.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 5.5, xmax = 6.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 7.5, xmax = 8.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 9.5, xmax = 10.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 11.5, xmax = 12.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +
  geom_rect(xmin = 13.5, xmax = 14.5, ymin = -Inf, ymax = 2.07, fill = "grey90", color = "gray90") +

  geom_abline(linetype = "dashed", slope = 0, intercept = 0.1) +
  geom_pointrange(data = subset(error_d15N, model %in% c("species","global","hette","jardine")),
                  aes(ymin = mean - sd, ymax = mean + sd),
                  position = position_dodge(width = 0.5),na.rm = TRUE) +
  geom_point(data = subset(error_d15N, model %in% c("species","global","hette","jardine")), 
             aes(x = factor(common_name, levels = rev(levels(factor(common_name)))), y = mean),
             position = position_dodge(width = 0.5), size = 3, color = "black", shape = 21, na.rm = TRUE) +
  geom_line(data = subset(error_d15N, model %in% c("species","global","hette","jardine")), 
            position = position_dodge(width = 0.5), na.rm = TRUE) + 
  coord_flip() +
  scale_y_continuous(limits = c(0,2)) +
  scale_colour_manual(values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF"), guide=FALSE) +
  scale_fill_manual(values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF"),
                      labels = c("Species specific", "Multi-species", "European","Australian")) +  
  labs(x = "", 
       y = "", 
       fill = "Conversion\nModel", 
       title = expression('(B)'~{delta}^15*N~'Error')) +
  theme_Publication()  +
  theme(axis.text.y = element_blank())
error_d15N_spp

p.error.panel <- 
  ggarrange(error_d13C_spp, error_d15N_spp, 
            widths = c(1,0.74), 
            common.legend = TRUE, 
            legend = "right"
            ) %>% 
  annotate_figure(bottom = text_grob(expression('Error '~{delta}^x*x~'\u2030'), 
                                     vjust = -1, hjust = 0.2))
p.error.panel

ggsave(filename = "figs/model-error-summary-panel.pdf",
       plot = p.error.panel, device = cairo_pdf,
       units = "in", width = 10, height = 6)

```

```{r Calculate error for each isotope and model}
error_d13C_test <- predictions_spp %>% 
  mutate(pred_global = predict(fit.global.C, data_test),
         pred_Hette = 0.82 * d13C_fin_lc - 5.89, 
         pred_Jardine  = 0.89 * d13C_fin_lc - 3.27, 
         
         species = abs(d13C_lc - pred_spp_C),
         global = abs(d13C_lc - pred_global),
         hette = abs(d13C_lc - pred_Hette), 
         jardine = abs(d13C_lc - pred_Jardine)) %>% 
  select(common_name, species, global, hette, jardine) %>%
  gather(key = "key", value = "value", -common_name) %>% 
  rename(model = key, mean = value) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(model = factor(model, levels = c("species","global", "hette","jardine"))) %>% 
  mutate(isotope = "Carbon") 
error_d13C_test

error_d15N_test <- 
  predictions_spp %>% 
  mutate(pred_global = predict(fit.global.N, data_test),
         pred_Hette = 1.01 * d15N_fin + 0.74, 
         pred_Jardine  = 0.81 * d15N_fin + 1.73, 
         
         species = abs(d15N - pred_spp_N),
         global = abs(d15N - pred_global),
         hette = abs(d15N - pred_Hette), 
         jardine = abs(d15N - pred_Jardine)) %>% 
  select(common_name, species, global, hette, jardine) %>%
  gather(key = "key", value = "value", -common_name) %>% 
  rename(model = key, mean = value) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(model = factor(model, levels = c("species","global", "hette","jardine"))) %>% 
  mutate(isotope = "Nitrogen")
error_d15N_test

error_df_test <- bind_rows(error_d13C_test, error_d15N_test)
error_df_test
```

```{r Error summary table}
error_summary <- 
  error_df_test %>% 
    group_by(isotope, model) %>% 
    summarise(mean_all = mean(mean), sd_all = sd(mean))  %>% 
    ungroup() %>% 
    mutate_if(is.numeric, round, 2) %>% 
  pivot_wider(names_from = isotope, values_from = c(mean_all, sd_all)) %>% 
    mutate(mean_err_C = paste0(mean_all_Carbon," (",sd_all_Carbon,")"), 
          mean_err_N = paste0(mean_all_Nitrogen," (",sd_all_Nitrogen,")")) %>% 
    select(-mean_all_Carbon, -mean_all_Nitrogen, -sd_all_Carbon, -sd_all_Nitrogen) %>% 
    arrange(model) %>% 
    write_csv("tables/error_summary.csv")
error_summary
```

## Test for differecnes amoung predicted model errors

```{r}
error_d13C_test <- predictions_spp %>% 
  mutate(pred_global = predict(fit.global.C, data_test),
         pred_Hette = 0.82 * d13C_fin_lc - 5.89, 
         pred_Jardine  = 0.89 * d13C_fin_lc - 3.27, 
         
         species = abs(d13C_lc - pred_spp_C),
         global = abs(d13C_lc - pred_global),
         hette = abs(d13C_lc - pred_Hette), 
         jardine = abs(d13C_lc - pred_Jardine)) %>% 
  select(common_name, species, global, hette, jardine) %>%
  gather(key = "key", value = "value", -common_name) %>% 
  rename(model = key, mean = value) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(model = factor(model, levels = c("species","global", "hette","jardine"))) 
error_d13C_test

error_d15N_test <- 
  predictions_spp %>% 
  mutate(pred_global = predict(fit.global.N, data_test),
         pred_Hette = 1.01 * d15N_fin + 0.74, 
         pred_Jardine  = 0.81 * d15N_fin + 1.73, 
         
         species = abs(d15N - pred_spp_N),
         global = abs(d15N - pred_global),
         hette = abs(d15N - pred_Hette), 
         jardine = abs(d15N - pred_Jardine)) %>% 
  select(common_name, species, global, hette, jardine) %>%
  gather(key = "key", value = "value", -common_name) %>% 
  rename(model = key, mean = value) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(model = factor(model, levels = c("species","global", "hette","jardine"))) 
error_d15N_test

```

```{r Stats}

hist(error_d13C_test$mean)
hist(error_d15N_test$mean)

shapiro_test(sqrt(error_d13C_test$mean))
shapiro_test(sqrt(error_d15N_test$mean))

hist(sqrt(error_d13C_test$mean))
hist(sqrt(error_d15N_test$mean))


error_d13C_test %>%
  mutate(mean = sqrt(mean)) %>% 
  anova_test(mean ~ model)

pairwise.carbon <- error_d13C_test %>% 
  mutate(mean = sqrt(mean)) %>% 
  t_test(mean ~ model, p.adjust.method = "bonferroni", detailed = TRUE) %>% 
  as_tibble() %>% 
  select(-n2, -p, -method, -alternative) %>% 
  mutate(y.position = c(1.4, 1.6, 2, 1.8, 2.2, 2.4))
pairwise.carbon

error_d13C_test %>%
  mutate(mean = sqrt(mean)) %>% 
  cohens_d(mean ~ model)



error_d15N_test %>%
  mutate(mean = sqrt(mean)) %>% 
  anova_test(mean ~ model)

pairwise.nitrogen <- error_d15N_test %>% 
  mutate(mean = sqrt(mean)) %>% 
  t_test(mean ~ model, p.adjust.method = "bonferroni", detailed = TRUE) %>% 
  as_tibble() %>% 
  select(-n2, -p, -method, -alternative) %>% 
  mutate(y.position = c(1.4, 1.6, 2, 1.8, 2.2, 2.4))
pairwise.nitrogen

error_d15N_test %>%
  mutate(mean = sqrt(mean)) %>% 
  cohens_d(mean ~ model)

```

```{r Stats}

hist(error_d13C$mean)
hist(error_d15N$mean)

shapiro_test(log10(error_d13C$mean))
shapiro_test(log10(error_d15N$mean))

shapiro_test(sqrt(error_d13C$mean))
shapiro_test(sqrt(error_d15N$mean))

hist(sqrt(error_d13C$mean))
hist(sqrt(error_d15N$mean))


error_d13C %>%
  mutate(mean = sqrt(mean)) %>% 
  anova_test(mean ~ model)

pairwise.carbon <- error_d13C %>% 
  mutate(mean = sqrt(mean)) %>% 
  t_test(mean ~ model, p.adjust.method = "bonferroni") %>% 
  as_tibble() %>% 
  mutate(y.position = c(1.4, 1.6, 2, 1.8, 2.2, 2.4))
pairwise.carbon

error_d15N %>%
  mutate(mean = sqrt(mean)) %>% 
  anova_test(mean ~ model)

pairwise.nitrogen <- error_d15N %>% 
  mutate(mean = sqrt(mean)) %>% 
  t_test(mean ~ model, p.adjust.method = "bonferroni") %>% 
  as_tibble() %>% 
  mutate(y.position = c(1.4, 1.6, 2, 1.8, 2.2, 2.4))
pairwise.nitrogen


```

```{r Error summarized boxplots}
p.error.c.box <- 
  error_d13C_test %>% 
  ggboxplot(x = "model", y = "mean", fill = "model",
            outlier.shape = NA, 
            palette =c("#440154FF", "#31688EFF", "#35B779FF", "#FDE725FF")) +
  geom_jitter(aes(fill = model), color="black", shape = 21,  size=1.5, alpha=0.5, 
              position=position_jitter(0.1)) +
  scale_x_discrete(labels = c("Species", "Multi-spp", "European","Australian")) +
  scale_y_continuous(limits = c(0,2.5)) +
  stat_summary(fun.y=mean, geom="point", shape=8, size=4, color="red", fill="red") +
  stat_compare_means(label.y = 2.5, size = 3, method = "anova") +
  stat_pvalue_manual(data = pairwise.carbon, label = "p.adj", size = 3,
                     xmin = "group1", xmax = "group2", y.position = "y.position") + 
  labs(x = "", y = expression('Error '~{delta}^13*C~'(\u2030)'), 
       title = expression('(A)'~{delta}^13*C~'Error')) +
  theme_Publication() +
  theme(legend.position = "none")

p.error.c.box

p.error.n.box <- 
  error_d15N %>% 
  ggboxplot(x = "model", y = "mean", fill = "model",
            outlier.shape = NA, 
            palette =c("#440154FF", "#31688EFF", "#35B779FF", "#FDE725FF")) +
  geom_jitter(aes(fill = model), color="black", shape = 21,  size=1.5, alpha=0.5, 
              position=position_jitter(0.1)) +
  scale_x_discrete(labels = c("Species", "Multi-spp", "European","Australian")) +
  scale_y_continuous(limits = c(0,2.5)) +
  stat_summary(fun.y=mean, geom="point", shape=8, size=4, color="red", fill="red") +
  stat_compare_means(label.y = 2.5, size = 3, method = "anova") +
  stat_pvalue_manual(data = pairwise.nitrogen, label = "p.adj", size = 3, 
                     xmin = "group1", xmax = "group2", y.position = "y.position") + 
  labs(x = "", y = expression('Error '~{delta}^15*N~'(\u2030)'), 
       title = expression('(B)'~{delta}^15*N~'Error')) +
  theme_Publication() +
  theme(legend.position = "none")
p.error.n.box

p.error.box.panel <- 
  ggarrange(p.error.c.box, p.error.n.box) %>% 
  annotate_figure(bottom = text_grob("Conversion Model", vjust = -1)) 
p.error.box.panel

ggsave(filename = "figs/fig4.pdf",
       plot = p.error.box.panel, device = cairo_pdf,
       units = "in", width = 8, height = 4)


```


# 6. Mixed effects models
ICC = intraclass correlation coeficient: proportion of total variance in y that is accounted for by group clustering. 

Random effects tells how much variance we find amoung levels of the grouping variable, plus the residual. 

The fixed effects is simlar to the lm. 

The random effect, species, is meant to capture all the influences of species on muscle isotope valules. 
```{r}
library(lme4)
library(lmerTest)
library(MuMIn)
library(sjPlot)
library(parameters)

```

```{r d15N model}
#===============================================================================
#### Multi-level analysis - Varying intercept model
#===============================================================================

m1 <- lmer(d15N ~ d15N_fin + (1 | taxon_code), data = data_train)  # allows untercepts to vary by g

summary(m1)
coef(m1) 
r.squaredGLMM(m1)  # get some R2 values, for fixed effects (R2m) and total (R2c)

#### plot the point estimates of intercept values with 95% CIs
sjPlot:::plot_model(m1, type = "re")

tab_model(m1)

#===============================================================================
#### Multi-level analysis - Varying intercept AND slope model
#===============================================================================

m2 <- lmer(d15N ~ d15N_fin + (1 + d15N_fin | taxon_code), data = data_train)
summary(m2)
round(0.94718/(0.94718 + 0.15233) * 100, digits = 0)

rand(m2)  # analysis of random effects - is variation associated with grouping significant? 
coef(m2)  # provides the final summation of fixed and random effects 
# That is, the addition of the random plus fixed effetcs
fixef(m2)  # global intercept and slope
ranef(m2)  # individual level alpha(j)'s; that is, the random effects
# add these alpha j's to the global alpha, and you get the coeficients
r.squaredGLMM(m2)  # get some R2 values, for fixed effects (R2m) and total (R2c)

coef(m2)$g
coef(m2)$g[1,]

fixef(m2)[1]
fixef(m2)[1] + ranef(m2)$taxon_code[1,1]  # calculation for intercept of that group
fixef(m2)[2]  # slope

ci(m2)
confint(m2)
standard_error(m2)
standard_error(m2, effects = "random")
model_parameters(m2, df_method = "satterthwaite")
coef(m2)

names <-enframe(rownames(coef(m2)$taxon_code))
x <- as_tibble(coef(m2)$taxon_code)
y <- as_tibble(standard_error(m2, effects = "random")$taxon_code)

bind_cols(names, x, y) %>% 
  rename(taxon_code = value, 
         intercept = `(Intercept)...3`, 
         slope = d15N_fin...4, 
         int_SE = `(Intercept)...5`, 
         slp_SE = d15N_fin...6) %>% 
  select(-name, taxon_code, intercept, int_SE, slope, slp_SE) %>% 
  left_join(data_fish_groups, by = "taxon_code") %>% 
  select(family, common_name, intercept, int_SE, slope, slp_SE) %>% 
  arrange(family, common_name) %>% 
  select(-family) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  write_csv("tables/mixed_coefs_nitorgen.csv")

#### plot the point estimates of intercept values with 95% CIs
sjPlot:::plot_model(m2, type = "std")
sjPlot:::plot_model(m2, type = "re")

tab_model(m2)

#### Compare the models #####
anova(m1, m2)
AIC(m1, m2, fit.global.C)

```

```{r d13C models}
#===============================================================================
#### Multi-level analysis - Varying intercept model
#===============================================================================

m3 <- lmer(d13C_lc ~ d13C_fin_lc + (1 | taxon_code), data = data_train)  # allows untercepts to vary by g

summary(m3)
round(0.1866/(0.1866 + 0.3005) * 100, digits = 0)

coef(m3) 
r.squaredGLMM(m3)  # get some R2 values, for fixed effects (R2m) and total (R2c)

#### plot the point estimates of intercept values with 95% CIs
sjPlot:::plot_model(m3, type = "re")

tab_model(m3)

#===============================================================================
#### Multi-level analysis - Varying intercept AND slope model
#===============================================================================

m4 <- lmer(d13C_lc ~ d13C_fin_lc + (1 + d13C_fin_lc | taxon_code), data = data_train)
summary(m4)
round(27.56770/(27.56770 + 0.25881) * 100, digits = 0)

confint(m4)
r.squaredGLMM(m4)  # much better than only varying intercept model

summary(m4)
# Fixed - global param values, significant effect of x on y
# Random effects - estimating vaiance ssociated with those g levels, and residual variance
rand(m4)  # analysis of random effects - is variation associated with grouping significant? 
coef(m4)  # provides the final summation of fixed and random effects 
# That is, the addition of the random plus fixed effetcs
fixef(m4)  # global intercept and slope
ranef(m4)  # individual level alpha(j)'s; that is, the random effects
# add these alpha j's to the global alpha, and you get the coeficients

names <-enframe(rownames(coef(m4)$taxon_code))
x <- as_tibble(coef(m4)$taxon_code)
y <- as_tibble(standard_error(m4, effects = "random")$taxon_code)

bind_cols(names, x, y) %>% 
  rename(taxon_code = value, 
         intercept = `(Intercept)...3`, 
         slope = d13C_fin_lc...4, 
         int_SE = `(Intercept)...5`, 
         slp_SE = d13C_fin_lc...6) %>% 
  select(-name, taxon_code, intercept, int_SE, slope, slp_SE) %>% 
  left_join(data_fish_groups, by = "taxon_code") %>% 
  select(family, common_name, intercept, int_SE, slope, slp_SE) %>% 
  arrange(family, common_name) %>% 
  select(-family) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  write_csv("tables/mixed_coefs_carbon.csv")

#### plot the point estimates of intercept values with 95% CIs
sjPlot:::plot_model(m4, type = "re")
sjPlot:::plot_model(m4, type = "pred")

tab_model(m3, m4)

#### Compare the models #####
anova(m3, m4)
AIC(m3, m4)
```

```{r}
tab_model(m4, m2)
```

```{r}
### plot
(mm_plot <- ggplot(data_train, aes(x = d13C_fin_lc, y = d13C_lc)) +
      lemon::facet_rep_wrap(~taxon_code) +   # a panel for each mountain range
      geom_point(alpha = 0.5) +
      theme_classic() +
      geom_line(data = cbind(data_train, pred = predict(m4)), aes(y = pred), size = 1) + 
      theme(legend.position = "none")
)
```

```{r}
library(ggeffects) 

# Extract the prediction data frame
pred.mm <- ggpredict(m4, terms = c("d13C_fin_lc"))  

# Plot the predictions 

(ggplot(pred.mm) + 
   geom_line(aes(x = x, y = predicted)) +          # slope
   geom_ribbon(aes(x = x, ymin = predicted - std.error, ymax = predicted + std.error), 
               fill = "lightgrey", alpha = 0.5) +  # error band
   geom_point(data = data_train,                      # adding the raw data (scaled values)
              aes(x = d13C_fin_lc, y = d13C_lc, colour = taxon_code)) + 
   labs(x = "fin", y = "muscle", 
        title = "Fins predict muscle values") + 
   theme_minimal()
)
```

```{r}
ggpredict(m2, terms = c("d15N_fin", "taxon_code"), type = "re") %>% 
   plot() +
   labs(x = "fin", y = "muscle", title = "Effect of species on muscle values") + 
   theme_minimal() + 
   scale_color_discrete()
```

# 7. Testing LMM models 

### Predictions
```{r Predict dataframe}
m2 <- lmer(d15N ~ d15N_fin + (1 + d15N_fin | taxon_code), data = data_train)
summary(m2)

# plot the marginal effects, not conditioned on random-effect variances.
pr <- ggpredict(m2, terms = "d15N_fin")
pr
plot(pr)

# Now account for random variance on popualtion estimate
pr <- ggpredict(m2, terms = "d15N_fin", type = "random")
pr
plot(pr)

# To get predicted values for a specific level of the random effect term, simply define this level in the condition-argument.
pr <- ggpredict(m2, terms = "d15N_fin", type = "random", condition = c(taxon_code = "CKC"))
pr
plot(pr)

# To compute marginal effects for each grouping level, add the related random term to the terms-argument. In this case, confidence intervals are not calculated, but marginal effects are conditioned on each group level of the random effects.
pr <- ggpredict(m2, terms = c("d15N_fin","taxon_code"), type = "random")
pr <- pr %>% filter(group %in% c("LNS","WHS","LND","CKC"))
plot(pr, add.data = TRUE)

pr <- ggpredict(m2, terms = c("d15N_fin","taxon_code [LNS, WHS, LND, CKC]"), type = "random")
plot(pr)

pr <- ggpredict(m2, terms = c("d15N_fin","taxon_code [sample=4]"), type = "random")
plot(pr)

pr <- ggpredict(m2, terms = c("d15N_fin","taxon_code [WHS]"), type = "random")
```
